Determination of Center of Rotation \\ in Computed Tomography

### 2017-07-22 seminar_cor/seminar_cor_haegele.tex
altered images, done abstract

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
The abstract goes here.
\end{abstract}

@@

This projection geometry can be mathematically expressed with the Radon transform.
When we think of the object as a function of attenuation coefficients $f(x,y)$ and the measured signal as a function of projected coefficients $R(r,\gamma)$ where $r$ is the position on the detector and $\gamma$ is the rotation angle at which the signal is obtained, then
\begin{equation}
\label{eq:radon}
R(r,\gamma) = \int_{-\infty}^{\infty} f(r\;cos(\gamma) + t\;sin(\gamma), r\;sin(\gamma)-t\;cos(\gamma))\;dt.
\end{equation} 
The parameter $t$ is the position on the ray that hits the detector at $r$ and angle $\gamma$, which makes it clear that the projection $R(r,\gamma)$ is the integral along a ray through the object $f(x,y)$.
The sampling of an object in this fashion, results in a so called sinogram.
The contribution of a specific area of the object to the individual projections, follows a sinusoidal path along the $\gamma$-axis as can be seen in \cref{fig:sinogram_parallel}
%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{img/phantomandradon.png}
\caption{
A \enquote{smilie} phantom and its corresponding sinogram with $\gamma~\in~[0^\circ,360^\circ]$.
An imaginary CT would have started with source on top and detector on bottom and then rotated clockwise around the phantom to obtain the shown sinogram.
}
\label{fig:sinogram_parallel}
\end{figure}
%

@@

\subsubsection{High Density Feature}\label{sssec:highdensityfeature}
Crawford~et~al.~\cite{crawford1988} presented an approach for finding the COR in which they used two $180^\circ$ opposing projections of a distinct high density feature.
When locating this feature in the two projections, the COR can be calculated as the sum of the detector coordinates of the feature divided by two.
This becomes clear when thinking about the feature being offset from the COR by $\delta$ in projection $p_1$.
The corresponding position of the feature in projection $p_2$ will then be on the opposite side of the COR, offset by $-\delta$ from the COR. \Cref{fig:highdenistyfeature} illustrates this concept.
%
\begin{figure}[!h]
\centering
\subfloat[phantom]{\includegraphics[width=0.32\linewidth]{img/highdensity_inv.png}}
\hfil
\subfloat[sinogram]{\includegraphics[width=0.65\linewidth]{img/highdensityradon_degrees_feature_centerline_horiz.png}}
\caption{
A phantom with a high density feature is shown on the left side.
In corresponding sinogram on the right, the sinusodial trace of the feature can be observed.
The dotted line indicates the COR in the sinogram.
It can be seen that the feature locations in $180^\circ$ opposing projections have the same distance to the COR.  
}
\label{fig:highdenistyfeature}
\end{figure}
%

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
This paper will discuss the aspect of the determination of center of rotation in computed tomography.
An introduction to computed tomography imaging will be given in the beginning as a prerequisite to the actual topic. 
This will cover different geometries (parallel beam, fan beam and cone beam) which are present in computed tomography and used to model the projection space of a CT-scan as well as basic knowledge about tomographic reconstruction, including the Fourier slice theorem and filtered back projection.
The problem of wrong center of rotation assumption in tomographic reconstruction will be explained which has the effect of image degradation.
Different solutions to the determination of the correct location of the center of rotation found in the literature will be presented.
It is found that the approaches can be split into sinogram and reconstruction based methods, where sinogram based methods analyze the sinogram that is the product of a CT-scan, while reconstruction based methods analyze reconstructed images of different centers of rotation to find the correct one.
Finally a conclusion summarizing the paper will be drawn.  
\end{abstract}

@@

This projection geometry can be mathematically expressed with the Radon transform.
When we think of the object as a function of attenuation coefficients $f(x,y)$ and the measured signal as a function of projected coefficients $R(r,\gamma)$ where $r$ is the position on the detector and $\gamma$ is the rotation angle at which the signal is obtained, then
\begin{equation}
\label{eq:radon}
R(r,\gamma) = \int_{-\infty}^{\infty} f(r\;cos(\gamma) + t\;sin(\gamma), r\;sin(\gamma)-t\;cos(\gamma))\;dt.
\end{equation} 
The parameter $t$ is the position on the ray that hits the detector at $r$ and angle $\gamma$, which makes it clear that the projection $R(r,\gamma)$ is the integral along a ray through the object $f(x,y)$.
The sampling of an object in this fashion, results in a so called sinogram.
The contribution of a specific area of the object to the individual projections, follows a sinusoidal path along the $\gamma$-axis as can be seen in \cref{fig:sinogram_parallel}
%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{img/phantomandradondarker.png}
\caption{
A \enquote{smilie} phantom and its corresponding sinogram with $\gamma~\in~[0^\circ,360^\circ]$.
An imaginary CT would have started with source on top and detector on bottom and then rotated clockwise around the phantom to obtain the shown sinogram.
}
\label{fig:sinogram_parallel}
\end{figure}
%

@@

\subsubsection{High Density Feature}\label{sssec:highdensityfeature}
Crawford~et~al.~\cite{crawford1988} presented an approach for finding the COR in which they used two $180^\circ$ opposing projections of a distinct high density feature.
When locating this feature in the two projections, the COR can be calculated as the sum of the detector coordinates of the feature divided by two.
This becomes clear when thinking about the feature being offset from the COR by $\delta$ in projection $p_1$.
The corresponding position of the feature in projection $p_2$ will then be on the opposite side of the COR, offset by $-\delta$ from the COR. \Cref{fig:highdenistyfeature} illustrates this concept.
%
\begin{figure}[!h]
\centering
\subfloat[phantom]{\includegraphics[width=0.32\linewidth]{img/highdensity_inv_darker.png}}
\hfil
\subfloat[sinogram]{\includegraphics[width=0.65\linewidth]{img/highdensityradon_degrees_feature_centerline_horiz.png}}
\caption{
A phantom with a high density feature is shown on the left side.
In corresponding sinogram on the right, the sinusodial trace of the feature can be observed.
The dotted line indicates the COR in the sinogram.
It can be seen that the feature locations in $180^\circ$ opposing projections have the same distance to the COR.  
}
\label{fig:highdenistyfeature}
\end{figure}
%
